---
title: "DS311 - R Lab Assignment"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
num_variables <- ncol(mtcars)
num_observations <- nrow(mtcars)
print(paste("There are a total of", num_variables, "variables and", num_observations, "observations in this data set."))

# Answer:
print("There are total of _____ variables and _____ observations in this data set.")
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary(mtcars)
num_discrete <- 3 
num_continuous <- 8
print(paste("There are", num_discrete, "discrete variables and", num_continuous, "continuous variables in this data set."))

# Answer:
print("There are _____ discrete variables and _____ continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
m <- mean(mtcars$mpg)
v <- var(mtcars$mpg)
s <- sd(mtcars$mpg)
print(paste("The average of Mile Per Gallon from this data set is", m, ", with variance", v, ", and standard deviation", s, "."))



# print(paste("The average of Mile Per Gallon from this data set is ",  , " with variance ",  , " and standard deviation",  , "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
average_mpg_cyl <- aggregate(mpg ~ cyl, data = mtcars, FUN = mean)
print("Average mpg for each cylinder class:")
print(average_mpg_cyl)

sd_mpg_gear <- aggregate(mpg ~ gear, data = mtcars, FUN = sd)
print("Standard deviation of mpg for each gear class:")
print(sd_mpg_gear)
```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
crosstab <- table(mtcars$cyl, mtcars$gear)
print("Crosstab of cylinder and gear class combinations:")
print(crosstab)

max_combination <- which.max(crosstab)
most_common_cyl <- as.numeric(rownames(crosstab)[max_combination %% nrow(crosstab)])
most_common_gear <- as.numeric(colnames(crosstab)[ceiling(max_combination / nrow(crosstab))])
num_most_common <- max(crosstab)

print(paste("The most common car type in this data set is car with", 
            most_common_cyl, "cylinders and", most_common_gear, 
            "gears. There are total of", num_most_common, 
            "cars belong to this specification in the data set."))

print("The most common car type in this data set is car with ____ cylinders and ____ gears. There are total of _____ cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
boxplot(weight ~ group, data = PlantGrowth,
        main = "Comparison of Plant Weights in Different Groups",
        xlab = "Group",
        ylab = "Weight of Plant",
        col = c("lightblue", "lightgreen", "lightpink"))

abline(h = median(PlantGrowth$weight), col = "red", lty = 2)

```

Result:

=> Report a paragraph to summarize your findings from the plot!
The boxplot shows a comparison of plant weights across three groups: control (ctrl), treatment 1 (trt1), and treatment 2 (trt2). The median plant weight for the control group is higher than that of treatment group 1, but lower than treatment group 2, suggesting that trt2 had the most positive effect on plant growth. Treatment 2 (trt2) had the highest median weight, indicating that it was the most effective treatment for promoting growth, with plants showing less variability in weight compared to other groups. In contrast, treatment 1 (trt1) had the lowest median weight, suggesting a less favorable outcome. Overall, treatment group 2 (trt2) produced the best results, while treatment group 1 showed the least effectiveness.


b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
hist(mtcars$mpg,
     breaks = 10,
     main = "Histogram of Miles per Gallon (mpg)",
     xlab = "Miles per Gallon",
     ylab = "Frequency",
     col = "lightblue")

hist_data <- hist(mtcars$mpg, breaks = 10, plot = FALSE)
most_frequent_mpg <- hist_data$mids[which.max(hist_data$counts)]

print(paste("Most of the cars in this data set are in the class of", most_frequent_mpg, "miles per gallon."))

```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
pairs(USArrests,
      main = "Pairs Plot for USArrests Data Set",
      col = "lightblue",
      pch = 19)

plot(USArrests$Murder, USArrests$Assault,
     main = "Scatter Plot: Murder vs Assault",
     xlab = "Murder Rate",
     ylab = "Assault Rate",
     pch = 19,
     col = "darkblue")

```

Result:

=> Report a paragraph to summarize your findings from the plot!
The findings from the USArrests dataset provide insights into the relationships between different crime variables across states. The pairs plot indicates positive correlations between Murder, Assault, UrbanPop, and Rape, with Murder and Assault exhibiting a particularly strong association. The scatter plot of Murder vs Assault further supports this observation, revealing a clear positive linear relationship, where states with higher murder rates also tend to have higher assault rates. This suggests that regions experiencing one form of violent crime are likely to report higher incidences of other violent crimes, highlighting potential underlying socio-economic or cultural factors that contribute to higher crime rates across the board.


***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set

# Enter your code here!
download.file(url = 'https://www.jaredlander.com/data/housing.csv', 
              destfile = 'C:/Users/KAUSTUBH/Desktop/Week_06_Import_Data_and_Visualization/housing.csv', 
              mode = 'wb')
housingData <- read.csv('C:/Users/KAUSTUBH/Desktop/Week_06_Import_Data_and_Visualization/housing.csv')
housingData <- subset(housingData, select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)

head(housingData)
summary(housingData)
aggregate(housingData$Market.Value.per.SqFt, by = list(Borough = housingData$Boro), FUN = mean)
aggregate(housingData$Year.Built, by = list(Neighborhood = housingData$Neighborhood), FUN = median)




## Question 3b - Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

# Enter your code here!
install.packages("ggplot2")
library(ggplot2)

ggplot(housingData, aes(x = Year.Built, y = Market.Value.per.SqFt)) +
  geom_point() +
  ggtitle("Market Value per SqFt vs Year Built") +
  xlab("Year Built") +
  ylab("Market Value per SqFt")

ggplot(housingData, aes(x = Boro, y = Market.Value.per.SqFt, color = Boro)) +
  geom_boxplot() +
  ggtitle("Market Value per SqFt by Borough") +
  xlab("Borough") +
  ylab("Market Value per SqFt")

pairs(~Market.Value.per.SqFt + Year.Built, data = housingData,
      main = "Scatterplot Matrix for Housing Data")

```

c. Write a summary about your findings from this exercise.

=> Enter your answer here!

First, the scatterplot showing "Market Value per SqFt vs Year Built" reveals a pattern indicating that older houses, particularly those built before the mid-1900s, tend to have a wide range of market values per square foot, whereas homes built more recently are concentrated in a narrower range. This suggests that homes built in different eras can have varying values based on market demand, location, and condition.

Additionally, the box plot "Market Value per SqFt by Borough" illustrates distinct differences among the boroughs in New York City. Manhattan has the highest median market value per square foot, followed by Brooklyn, while Staten Island and the Bronx have lower market values. This suggests that location plays a significant role in determining property value. The range of property values within Manhattan also demonstrates high variability, likely influenced by different neighborhoods or unique property attributes.

The summary statistics from the dataset indicate that neighborhoods like Tribeca and the Financial District in Manhattan have higher market values compared to others, further emphasizing Manhattan's premium in housing prices.
